Public data: domestic dogs

### 2018-01-18 GWAS-comparison.Rmd
save

---
title: "Comparing bigstatsr and bigsnpr with PLINK and EIGENSOFT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7)
options(width = 85)
```

@@

Here, we suppose that the data have been pre-processed (quality control and imputation), and is available as a PLINK binary file.

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# FastPCA: https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz
smartpca <- "../EIG-6.1.4/bin/smartpca"
plink <- download_plink()

@@

In this section, we will show how to perform a GWAS with or without our packages. For simplicity, everything will be done inside R.

@@

```{r}
celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos

@@

ggsave("figures/celiac-pca.png", width = 1124, height = 945, scale = 1/75)
```

@@

ggsave("figures/celiac-gwas-cut.png", width = 1475, height = 788, scale = 1/75)
```

@@

```{r}
# Pruning
snp_writeBed(celiac, bedfile = "plink0.bed")
prefix_bed <- "plink0"
write.table2(LD.wiki34, "LRLDR.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --exclude LRLDR.txt --range",
              " --indep-pairwise 50 1 0.2",
              " --threads {NCORES}"))
)

ind.keep2 <- match(scan("plink.prune.in", what = ""), celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

# Make a pruned dataset
write.table2(as.data.frame(celiac$map[ind.keep, "marker.ID"]), "keep.in")
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --extract keep.in --make-bed",
              " --threads {NCORES}"))
)

# PCA with EIGENSOFT
fastpca <- function(prefix, fast = TRUE) {
  parfile <- glue("{prefix}.pca.par")

@@

system.time(
  obj.svd2 <- fastpca("plink")
)

# FastPCA is not accurate enough (see benchmark-pca.html)
round(100 * cor(obj.svd2, obj.svd$u), 1)
```

```{r}
# GWAS
write.table2(cbind(celiac$fam[, 1:2], obj.svd$u), "covar.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --logistic hide-covar",
              " --covar covar.txt",
              " --threads {NCORES}"))
)

gwas2 <- fread2("plink.assoc.logistic")
all.equal(gwas2$P, predict(gwas, log10 = FALSE)) 
```

## Example of Polygenic Risk Scores (PRS)

In this section, we will show how to create a PRS with or without our packages. For simplicity, everything will be done inside R.

```{r}
# Split in training/test set (80%/20%)
ind.train <- sort(sample(nrow(G), 0.8 * nrow(G)))
ind.test <- setdiff(rows_along(G), ind.train)
```

### With bigstatsr and bigsnpr

```{r}
# GWAS on training set
y01 <- celiac$fam$affection - 1
system.time(
  gwas.train <- big_univLogReg(G, y01.train = y01[ind.train],
                               ind.train = ind.train, 
                               covar.train = obj.svd$u[ind.train, ], 
                               ncores = NCORES)
)

# Clumping
system.time(
  ind.keep <- snp_clumping(G, infos.chr = celiac$map$chromosome,
                           ind.row = ind.train,
                           S = abs(gwas.train$score),
                           size = 250, # as PLINK default
                           is.size.in.bp = TRUE,
                           infos.pos = celiac$map$physical.pos,
                           ncores = NCORES)
)

# PRS
(thrs <- c(0, exp(seq(log(0.1), log(100), length.out = 100))))
lpS <- -predict(gwas.train)
system.time(
  prs <- snp_PRS(G, betas.keep = gwas.train$estim[ind.keep],
                 ind.test = ind.test,
                 ind.keep = ind.keep,
                 lpS.keep = lpS[ind.keep], 
                 thr.list = thrs)
)

# AUC
aucs <- apply(prs, 2, AUC, target = y01[ind.test])
# Number of predictor for each threshold
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
```

### With PLINK (and R)

```{r, fig.width = 10}
# Get a file with only training set
write.table2(cbind(celiac$fam[ind.train, 1:2], 
                   obj.svd$u[ind.train, ]), "covar_train.txt")
# GWAS on training set
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --keep covar_train.txt",
              " --logistic hide-covar beta",
              " --covar covar_train.txt",
              " --threads {NCORES}"))
)

gwas2 <- fread2("plink.assoc.logistic")
all.equal(gwas2$P, 10^predict(gwas.train))

# Clumping
system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --keep covar_train.txt",
              " --clump plink.assoc.logistic",
              " --clump-p1 1 --clump-p2 1 --clump-r2 0.2",
              " --threads {NCORES}"))
)

clump <- fread2("plink.clumped")
ind.keep2 <- match(clump$SNP, celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

gwas2.keep <- gwas2[ind.keep, ]
write.table2(gwas2.keep, "assoc_clumped.txt")

# PRS
write.table2(celiac$fam[ind.test, 1:2], "ind_test.txt")
write.table2(cbind(paste0("S", seq_along(thrs)), 0, 10^(-thrs)),
             "range.txt")

system.time(
  system(glue("{plink} --bfile {prefix_bed}",
              " --keep ind_test.txt",
              " --extract assoc_clumped.txt",
              " --score assoc_clumped.txt 2 4 7 sum",
              " --q-score-range range.txt assoc_clumped.txt 2 9",
              " --threads {NCORES}"))
)

scores2 <- matrix(NA_real_, length(ind.test), length(thrs))
nb.pred2 <- numeric(length(thrs))
for (i in seq_along(thrs)) {
  tmp <- fread2(glue("plink.S{i}.profile"))
  scores2[, i] <- tmp$SCORESUM
  nb.pred2[i]  <- tmp$CNT[[1]] / 2
}

#AUC
aucs2 <- apply(scores2, 2, AUC, target = y01[ind.test])

bind_rows(
  data_frame(nb = nb.pred,  auc = aucs,  Software = "bigsnpr"),
  data_frame(nb = nb.pred2, auc = aucs2, Software = "PLINK")
) %>%
  ggplot() %>%
  bigstatsr:::MY_THEME() +
  geom_point(aes(nb, auc, color = Software, shape = Software), size = 5) +
  geom_line(aes(nb, auc, color = Software, linetype = Software), show.legend = FALSE) +
  scale_x_log10() + 
  scale_colour_manual(
    values = c("black", "red"),
    labels = c("PRS with bigstastr\nand bigsnpr", "PRS with plink")) + 
  scale_shape_manual(
    values = c(1, 3),
    labels = c("PRS with bigstastr\nand bigsnpr", "PRS with plink")) +
  labs(x = "#predictors", y = "AUC") +
  theme(legend.position = c(0.6, 0.2)) 
```

```{r}
# Cleaning
file.remove(c(list.files(pattern = "plink*|*.txt"), "keep.in"))
```

>>>>

---
title: "Genome-Wide Association Study"
subtitle: "Comparing bigstatsr and bigsnpr with PLINK and EIGENSOFT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7)
options(width = 90)
```

@@

Here, we suppose that the data have been pre-processed (quality control and imputation), and is available as a bk/rds (format of packages bigstatsr and bigsnpr). See the notebook "preprocessing".

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# download.file("https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz",
#               destfile = "EIG-6.1.4.tar.gz")
# untar("EIG-6.1.4.tar.gz")
smartpca <- "EIG-6.1.4/bin/smartpca"
plink <- download_plink()

@@

We will show how to perform a GWAS with or without our packages. For the sake of reproducibility and simplicity, everything will be done inside R (via possible system calls to command-line software).

@@

```{r}
system.time(
  celiac <- snp_attach("backingfiles/celiacQC.rds")
)
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos

@@

ggsave("figures/celiac-pca.png", width = 1124, height = 945, scale = 1/90)
```

@@

ggsave("figures/celiac-gwas-cut.png", width = 1475, height = 788, scale = 1/90)
```

@@

```{r}
# Pruning
prefix_bed <- tempfile()
snp_writeBed(celiac, bedfile = paste0(prefix_bed, ".bed"))
write.table2(LD.wiki34, "LRLDR.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --exclude LRLDR.txt --range",
              " --indep-pairwise 50 1 0.2",
              " --threads {NCORES}"))
)

ind.keep2 <- match(scan(paste0(prefix_bed, ".prune.in"), what = ""), celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

# Make a pruned dataset
write.table2(as.data.frame(celiac$map[ind.keep, "marker.ID"]), "keep.in")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --extract keep.in --make-bed",
              " --threads {NCORES}"))
)

# PCA with EIGENSOFT
fastpca <- function(prefix, fast = TRUE) {

  parfile <- glue("{prefix}.pca.par")

@@

system.time(
  obj.svd2 <- fastpca(tmp)
)

# FastPCA is not accurate enough (see Notebook "benchmark-pca")
round(100 * cor(obj.svd2, obj.svd$u), 1)
```

```{r}
# GWAS
write.table2(cbind(celiac$fam[, 1:2], obj.svd$u), "covar.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --logistic hide-covar",
              " --covar covar.txt",
              " --threads {NCORES}"))
)

gwas2 <- fread2(paste0(prefix_bed, ".assoc.logistic"))
all.equal(gwas2$P, predict(gwas, log10 = FALSE)) 
```

### 2018-01-18 cabios-template2/bigsnpr-review.tex
save

Consider first the basic tools necessary to perform a standard genome-wide analysis. Conversions between standard file formats has become a field by itself with several tools such as VCFtools, BCFtools and PLINK, available either independently or incorporated within large framework \cite[]{Danecek2011,Li2011,Purcell2007}. Similarly, quality control software for genome-wide analysis have been developed such as PLINK and the Bioconductor package GWASTools \cite[]{Gogarten2012}.
There are also several software for the computation of principal components (PCs) of genotypes, commonly performed to account for population stratification in association studies, including EIGENSOFT (SmartPCA and FastPCA) and FlashPCA \cite[]{Abraham2014a,Abraham2016a,Galinsky2016,Price2006}.
Then, implementation of GWAS analyses also depends on the data format and model analyzed. For example, the software ProbABEL \cite[]{Aulchenko2010} or SNPTEST \cite[]{Marchini2010} can handle dosage data, while PLINK version 1.9 is limited to best guess genotypes because of its input file format. Note that PLINK 2.0 will completely handle this type of data, but is still in the alpha stage of development. 
Finally, there exists a range of tools for Polygenic Risk Scores (PRSs) such as LDpred \cite[]{Vilhjalmsson2015} and PRSice \cite[]{Euesden2015}, which provide prediction for quantitative traits or disease risks based on multiple genetic variants. As a result, one has to make extensive bash/perl/R/python scripts to link these software together and convert between multiple file formats, involving many file manipulations and conversions. Overall, this might be a barrier to data exploration. 
%Overall, this means that researchers are usually restricted on how they can manipulate and analyse the data they have access to.

@@

Most modern SNP chips provide genotype data with large call-rates. For example, the Celiac data we use in this paper present only a proportion of 0.0004 of missing values for the genotyped SNPs after quality control. Yet, most of the functions in bigstatsr and bigsnpr do not handle missing values.
So, we provide two functions for imputing missing values of genotyped SNPs. Note that we do not impute completely missing SNPs which would require the use of reference panels and could be performed via e.g. imputation servers for human data \cite[]{mccarthy2016reference}. The first function is a wrapper to PLINK and Beagle \cite[]{Browning2009} which takes bed files as input and return bed files without missing values, and should therefore be used before reading the data in R (Figure~\ref{fig:impute}). Beagle uses phasing before imputing, which could take several weeks even when there are only a tiny amount of missing values to be imputed. The second function is a new algorithm we developed in order to have a fast imputation method without losing much of imputation accuracy. We also provide an accurate estimation of the number of errors of imputation by SNP for post quality control.
This algorithm is based on Machine Learning approaches for genetic imputation \cite[]{Wang2012} and does not use phasing, thus allowing for a dramatic decrease in computation time. It only relies on some local XGBoost models. XGBoost is an optimized distributed gradient boosting library that can be used in R and provides some of the best results in machine learning competitions \cite[]{Chen2016}.
XGBoost builds decision trees that can detect nonlinear interactions, partially reconstructing phase, making it well suited for imputing genotype matrices. 
Our algorithm is the following: for each SNP, we divide the individuals in the ones which have a missing genotype (test set) and the ones which have a non-missing genotype for this particular SNP. Those latter individuals are further separated in a training set and a validation set (e.g. 80\% training and 20\% validation). The training set is used to build the XGBoost model for predicting missing data. The prediction model is then evaluated on the validation set for which we know the true genotype values, which gives an unbiased estimator of the number of genotypes that have been wrongly imputed for that particular SNP. The prediction model is also projected on the test set (missing values) in order to impute them.
}

@@

SNP thinning improves ascertainment of population structure with PCA \cite[]{Abdellaoui2013}. There are at least 3 different approaches to thin SNPs based on Linkage Disequilibrium, two of them named pruning and clumping, address SNPs in LD close to each others because of recombination events, while the third one address long-range regions with a complex LD pattern due to other biological events such as inversions \cite[]{Price2008}. 
First, pruning, the most naive approach, is an algorithm that sequentially scan the genome for nearby SNPs in LD, performing pairwise thinning based on a given threshold of correlation.
A variant of pruning is clumping. 
{\color{red}
Clumping is useful if a statistic is available to sort the SNPs by importance. Clumping is usually used to post-process results of a GWAS in order to keep only the most significant SNP per region of the genome. For PCA, the thinning procedure should remain unsupervised (one must not use any phenotype) and we therefore propose to use the minor allele frequency (MAF) as the statistic of importance. Note that this choice does not come out of nowhere because in the pruning algorithm of PLINK, when two nearby SNPs are correlated, only the one with the higher MAF is kept. Yet, in some worst-case scenario, the pruning algorithm can leave regions of the genome without any representative SNP at all so we advise to always use clumping instead of pruning, using the MAF as the statistic of importance, which is the default in function snp\_clumping ([SUPMAT]).
}

@@

In this paper, two datasets are analyzed: the celiac disease cohort and the POPRES datasets \cite[]{Dubois2010,Nelson2008}. The Celiac dataset is composed of 15,283 individuals of European ancestry genotyped on 295,453 SNPs. The POPRES dataset is composed of 1385 individuals of European ancestry genotyped on 447,245 SNPs.
For computation times comparison, we replicated individuals in the Celiac dataset 5 and 10 times in order to increase sample size while keeping the same {\color{red} eigen decomposition (up to a constant) and pairwise SNP correlation} as the original dataset. To assess scalibility of our algorithms for a biobank-scale genotype dataset, we formed another dataset of 500,000 individuals and 1 million SNPs, also through replication of the Celiac dataset. \label{sec:rep}

@@

All the code used in this paper along with results, such as execution times and figures, are available as HTML R notebooks in the Supplementary Data.

>>>>

Consider first the basic tools necessary to perform a standard genome-wide analysis. Conversions between standard file formats has become a field by itself with several tools such as VCFtools, BCFtools and PLINK, available either independently or incorporated within large framework \cite[]{Danecek2011,Li2011,Purcell2007}. Similarly, quality control software for genome-wide analysis have been developed such as PLINK and the Bioconductor package GWASTools \cite[]{Gogarten2012}.
There are also several software for the computation of principal components (PCs) of genotypes, commonly performed to account for population stratification in association studies, including EIGENSOFT (SmartPCA and FastPCA) and FlashPCA \cite[]{Abraham2014a,Abraham2016a,Galinsky2016,Price2006}.
Then, implementation of GWAS analyses also depends on the data format and model analyzed. For example, the software ProbABEL \cite[]{Aulchenko2010} or SNPTEST \cite[]{Marchini2010} can handle dosage data, 
{\color{red}
while PLINK version 1.9 is limited to best guess genotypes because of its input file format. Note that PLINK 2.0 will completely handle this type of data, but is still in the alpha stage of development. 
}
Finally, there exists a range of tools for Polygenic Risk Scores (PRSs) such as LDpred \cite[]{Vilhjalmsson2015} and PRSice \cite[]{Euesden2015}, which provide prediction for quantitative traits or disease risks based on multiple genetic variants. As a result, one has to make extensive bash/perl/R/python scripts to link these software together and convert between multiple file formats, involving many file manipulations and conversions. Overall, this might be a barrier to data exploration. 
%Overall, this means that researchers are usually restricted on how they can manipulate and analyse the data they have access to.

@@

Most modern SNP chips provide genotype data with large call-rates. For example, the Celiac data we use in this paper presents only a proportion of 0.0004 of missing values for the genotyped SNPs after quality control. Yet, most of the functions in bigstatsr and bigsnpr do not handle missing values.
So, we provide two functions for imputing missing values of genotyped SNPs. Note that we do not impute completely missing SNPs which would require the use of reference panels and could be performed via e.g. imputation servers for human data \cite[]{mccarthy2016reference}. The first function is a wrapper to PLINK and Beagle \cite[]{Browning2009} which takes bed files as input and return bed files without missing values, and should therefore be used before reading the data in R (Figure~\ref{fig:impute}). Beagle uses phasing before imputing, which could take several weeks even when there are only a tiny amount of missing values to be imputed. The second function is a new algorithm we developed in order to have a fast imputation method without losing much of imputation accuracy. We also provide an accurate estimation of the number of errors of imputation by SNP for post quality control.
This algorithm is based on Machine Learning approaches for genetic imputation \cite[]{Wang2012} and does not use phasing, thus allowing for a dramatic decrease in computation time. It only relies on some local XGBoost models. XGBoost is an optimized distributed gradient boosting library that can be used in R and provides some of the best results in machine learning competitions \cite[]{Chen2016}.
XGBoost builds decision trees that can detect nonlinear interactions, partially reconstructing phase, making it well suited for imputing genotype matrices. 
Our algorithm is the following: for each SNP, we divide the individuals in the ones which have a missing genotype (test set) and the ones which have a non-missing genotype for this particular SNP. Those latter individuals are further separated in a training set and a validation set (e.g. 80\% training and 20\% validation). The training set is used to build the XGBoost model for predicting missing data. The prediction model is then evaluated on the validation set for which we know the true genotype values, which gives an unbiased estimator of the number of genotypes that have been wrongly imputed for that particular SNP. The prediction model is also projected on the test set (missing values) in order to impute them.
}

@@

SNP thinning improves ascertainment of population structure with PCA \cite[]{Abdellaoui2013}. There are at least 3 different approaches to thin SNPs based on Linkage Disequilibrium, two of them named pruning and clumping, address SNPs in LD close to each others because of recombination events, while the third one address long-range regions with a complex LD pattern due to other biological events such as inversions \cite[]{Price2008}. 
First, pruning, the most naive approach, is an algorithm that sequentially scan the genome for nearby SNPs in LD, performing pairwise thinning based on a given threshold of correlation.
A variant of pruning is clumping. 
{\color{red}
Clumping is useful if a statistic is available to sort the SNPs by importance. Clumping is usually used to post-process results of a GWAS in order to keep only the most significant SNP per region of the genome. For PCA, the thinning procedure should remain unsupervised (one must not use any phenotype) and we therefore propose to use the minor allele frequency (MAF) as the statistic of importance. Note that this choice does not come out of nowhere because in the pruning algorithm of PLINK, when two nearby SNPs are correlated, only the one with the higher MAF is kept. Yet, in some worst-case scenario, the pruning algorithm can leave regions of the genome without any representative SNP at all ([SUPMAT]). So, we suggest to always use clumping instead of pruning, using the MAF as the statistic of importance, which is the default in function snp\_clumping . In practice, we tested the effect of using clumping instead of pruning for the thinning step before PCA and the results were very similar.
}

@@

In this paper, two datasets are analyzed: the celiac disease cohort and the POPRES datasets \cite[]{Dubois2010,Nelson2008}. The Celiac dataset is composed of 15,283 individuals of European ancestry genotyped on 295,453 SNPs. The POPRES dataset is composed of 1385 individuals of European ancestry genotyped on 447,245 SNPs.
For computation times comparison, we replicated individuals in the Celiac dataset 5 and 10 times in order to increase sample size while keeping the same {\color{red} eigen decomposition (up to a constant) and pairwise SNP correlations} as the original dataset. To assess scalibility of our algorithms for a biobank-scale genotype dataset, we formed another dataset of 500,000 individuals and 1 million SNPs, also through replication of the Celiac dataset. \label{sec:rep}

@@

All the code used in this paper along with results, such as execution times and figures, are available as HTML R notebooks in the Supplementary Data. 
{\color{red}
We now provide some open-access data of domestic dogs so that users can test our code and functions on a moderate size dataset with 4342 samples and 145,596 SNPs ([SUPMAT]) \cite[]{hayward2016complex}.
}

### 2018-01-18 preprocessing.Rmd
save

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

NCORES <- nb_cores()

popresQC.bed <- snp_plinkQC(
  prefix.in = "../POPRES_data/POPRES_allchr",
  file.type = "--file", # ped/map
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

popresQC2.bed <- snp_plinkIBDQC(popresQC.bed, ncores = NCORES)
```

```{r}
celiacQC.bed <- snp_plinkQC(
  prefix.in = "../thesis-celiac/Dubois2010_data/FinnuncorrNLITUK1UK3hap300",
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

celiacQC2.bed <- snp_plinkIBDQC(celiacQC.bed, ncores = NCORES)
```

@@

```{r}
popresQC.rds <- snp_readBed(popresQC2.bed, "backingfiles/popresQC")

@@

```{r}
system.time(
  celiac <- snp_attach(celiacQC.rds)
)
str(celiac, max.level = 2)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7)
options(width = 90)
```

@@

plink <- download_plink()
NCORES <- nb_cores()

popresQC.bed <- snp_plinkQC(
  prefix.in = "data/POPRES_allchr",
  plink.path = plink,
  file.type = "--file", # ped/map
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

popresQC2.bed <- snp_plinkIBDQC(
  bedfile.in = popresQC.bed, 
  plink.path = plink, 
  ncores = NCORES
)
```

```{r}
celiacQC.bed <- snp_plinkQC(
  prefix.in = "data/FinnuncorrNLITUK1UK3hap300",
  plink.path = plink,
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

celiacQC2.bed <- snp_plinkIBDQC(
  celiacQC.bed, 
  plink.path = plink,
  ncores = NCORES
)
```

@@

```{r}
system.time(
  popresQC.rds <- snp_readBed(popresQC2.bed, "backingfiles/popresQC")
)

@@

```{r}
celiac <- snp_attach(celiacQC.rds)
str(celiac, max.level = 2)
```

@@

The above plot is useful for post-processing the imputation.
No SNP (point) is above the red line. This means that no SNP has more than 1% of error.

### 2018-01-18 public-data.Rmd
save

```{r}
system.time(
  rds <- snp_readBed("backingfiles/cornell_canine_QC.bed", "backingfiles/dogs")
)
system.time(
  bigsnp <- snp_attach(rds)
)
```

@@

```{r}

```

>>>>

```{r}
system.time(
  rds <- snp_readBed("backingfiles/cornell_canine_QC.bed", "backingfiles/dogs")
)
system.time(
  bigsnp <- snp_attach("backingfiles/dogs.rds")
)
```

@@
